Tropical Mathematics for the Lambda Calculus

### 2023-01-16 section3A.tex
Proposal commit
Commit sul linear lam-calc

@@

\begin{fact}\label{fact:LREL_SMCC}
 $\LREL$ is a Symmetric Monoidal Closed Category.
\end{fact}

Its monoidal structures is given by a tensor product $\otimes$ acting on the objects as the Cartesian product of sets, and as the Kronecker product of matrices on morphisms.
Its closed structures is given by $X\multimap Y:=X\times Y$ with the usual evaluation and curry morphisms yielding a symmetric monoidal adjunction: $\LREL(Z\otimes X, Y) \simeq \LREL (Z, X\multimap Y)$.

Fact~\ref{fact:LREL_SMCC} means that $\LREL$ is a model of Intuitionistic Multiplicative Linear Logic (IMLL), and thus we can interpret \emph{linear} $\lam$-terms in it.

>>>>

The linear simply typed $\lam$-calculus is a restriction of the ordinary $\lam$-calculus in which programs can only use their arguments \emph{exatly once} during the computation.
This is the notion of linearity taken into account by linear logic, and indeed this calculus can be embedded in the Intuitionistic Multiplicative fragment of it (called \emph{IMLL}).

More precisely, the linear $\lam$-calculus is obtained from the ordinary one by adding the constraint that each $\lam$-abstracted variable appears exactly once in the scope of the $\lam$-abstraction.

In order to frame it in a category $\C C$, one needs a symmetric tensor product $\otimes$ together with exponential objects $X\multimap Y$ s.t.\ the evaluation and curry maps yield a \emph{symmetric monoidal adjunction}: $\C C(Z\otimes X, Y) \simeq \C C (Z, X\multimap Y)$.
This is the notion of Symmetric Monoidal Closed Category (SMCC).

@@

\begin{fact}\label{fact:LREL_SMCC}
 $\LREL$ is a SMCC, thus a model of the linear $\STLC$.
\end{fact}

Its monoidal structures is given by a tensor product $\otimes$ acting on the objects as the Cartesian product of sets, and as the \emph{Kronecker product} {(\color{red} Ref?)} of matrices on morphisms.
Its closed structures is given by $X\multimap Y:=X\times Y$ with the usual evaluation and curry maps.
Actually, the SMCC $\LREL$ is even a model of IMLL.
